### 2007-12-19 trunk/paper/paper.tex
initial port from -r372 of the GoogleDoc version

\documentclass{memoir}

\usepackage{hyperref}
\usepackage{memhfixc}

\title{Jaigo Technical Report}
\author{Sean Foy \and Noel Weichbrodt}

\begin{document}

\begin{titlingpage}
\maketitle

\begin{abstract}
Despite its elegantly simple rules, Go's complexity as a game of deterministic strategy has made it a recurring subject in studies of artificial intelligence and game theory \cite{Muller2002}.
\end{abstract}
\end{titlingpage}

\tableofcontents
\listoftables

\chapter{Introduction}
%Include background materials and define the problem and scope of your project.
Go is a two-player board game which originated in Ancient China and remains one of the most richly complex games of strategy today. The objective is to control more area or territory than one's opponent on the game board. The game is played on a square line grid; the standard size of this grid is 19 x 19, though 13 x 13 and 9 x 9 grids are used occasionally. Players place stones -- either black or white -- at the intersecting points of the grid to signify control of that position. Once placed on the board a stone has a number of liberties -- unoccupied grid positions which are orthogonally adjacent to that stone. The goal of the game is to claim as much territory as possible by forming chains (contiguous extensions of stones of the same color) which cannot be captured by one's opponent. A stone or chain is captured and removed from the board when it has no liberties, the result of being surrounded by stones of the opposing color.

Players are not allowed to commit ``suicide'' with their stones, that is, they may not play into a spot where a stone would have no liberties. This prohibition actually enables players to create certain arrangements of stones which cannot be captured. Players are also forbidden from repeating one of the previous two states of the game; players cannot repeat their previous move. This is known as the Ko rule. At each turn, a player can either place a stone or pass. The game ends when both players pass consecutively on a turn.

Go players are classified by their skill into equivalence classes called kyu and dan. A beginning player might have a rank of 30 kyu, while a more experienced player might be 15kyu. Avid amateurs and professionals who outrank 1 kyu fill the range from 1 dan (low) to 8 dan (high) and on to 1 professional dan (low) to 10p (high).

Game theorists would describe Go as a zero-sum, perfect information, partisan, deterministic strategy game. In zero-sum games, the utility of any action with respect to player a is the additive inverse of the utility of that action with respect to player b. Perfect information games are fully observable in AI parlance. Go has a turn-taking structure, so it is partisan (i.e., some moves are available to only one player). There is no element of chance in Go. These characteristics invite comparison with games such as Chess.

While chess eventually yielded to clever hardware and algorithms, the sheer state-space of Go has rendered specialize search approaches weak. In 1965, noted statistician I.J. Good noted presciently:

\begin{quote}
Go on a computer? – In order to programme [sic] a computer to play a reasonable game of Go – rather than merely a legal game – it is necessary to formalise the principles of good strategy, or to design a learning programme. The principles are more qualitative and mysterious than in chess, and depend more on judgment. So I think it will be even more difficult to programme a computer to play a reasonable game of Go than of chess. \cite{Good65}
\end{quote}

Unlike many similar problems that were also announced in this manner, Go remains both a challenging and tantalizing problem.

Two major factors make computer Go more difficult to program than computer Chess:
\begin{enumerate}
  \item Go is PSPACE-hard with a branching factor of 361, and so is not amenable to brute force searching.
  \item The traditional concepts and representations of the state of the board do not translate easily to traditional areas of AI research.
\end{enumerate}

\chapter{Related theory and practice}
\section{Board representation}
With 3361 board states, and perhaps 10768 games, representing the board may be the most difficult part of writing a go engine.

The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memoize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}. The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.

\section{Known techniques}
Search techniques have been successful in games including Chess, but they have been much less effective in Go. The large branching factor of Go prohibits exhaustive search. Wolfe has shown by reduction from 3-QBF that Go endgames are PSPACE-hard \cite{Wolfe} and Go itself is EXPTIME-complete \cite{Burmeister}, so it is unlikely that any algorithm can guarantee optimal and efficient play. Computer Go is thus focused on approximation algorithms, and suboptimal or incomplete but empirically successful heuristic searches. Minimax with $\alpha$-$\beta$ pruning and the anytime beam stack search methods are of limited use given the very short horizon and absence of good heuristics for Go.

The difficulty of formalizing the heuristics used by humans motivates us to consider applying machine learning algorithms to pattern-matching in Go. Researchers including \cite{Enzenberger96} have enjoyed limited success with neural networks, but both training and playing are slow, and feature vector characterization is difficult. Strategies for overcoming the need for a priori knowledge in feature vector formulation and connection weights include constructive and genetic algorithms. Unfortunately, experiments with these techniques have yielded ``erratic and unorganized'' play despite ``prohibitively large'' time requirements for playing on full-size boards \cite{RichardsMM98}.

Br\"ugmann's approach plays games by choosing moves randomly according to probability distributions found by simulated annealing; the moves are evaluated according to their occurrence in successful games. A key insight from this research is that moves can be iteratively reprioritized by observed value rather than played in some fixed order; this is crucial for the simulated annealing application \cite{Bruegmann93}.

\chapter{Current Issues}

    * Finding a good evaluation function is hard. The human heuristics are vague, and when/how they get applied is even more vague. Zobrist divided the game into phases for the purpose of selecting evaluation functions \cite{Zobrist1970b}. Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function). GNUGO v Go++
    * GA \& NN seemed promising at first, but no commercial package uses them due to lack of speed.
    * If all old techniques are inadequate and new techniques don't pan out, whither computer go?

Our project seeks to create a JavaScript-based Go-engine. JavaScript is a dynamically typed, imperative language that supports the functional and object-oriented paradigms \cite{ECMA-262}. It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail. As professional programmers, we viewed this project as an opportunity to explore not only AI but also renewed interest in the JavaScript language. Also, JavaScript is the only supported means of developing iPhone applications. iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations only heighten the interest.

\chapter{Interaction/UI}

We decided that our software should be interoperable with other Go software, allowing us to focus on the AI and making our software more useful. The Go Text Protocol \cite{GTP} is a defacto standard for interoperability, so we have focused our effort on implementing the role of the engine in the GTP. Besides providing an interface to competitive engines and automated testing frameworks, it allows us to avoid writing our own UI by reusing any of several existing UI implementations. During early development, we used a simple set of HTML input controls to interact with the engine via GTP.

\chapter{Scoring}

The board is scored using Benson's Algorithm \cite{Benson76}. ``c'' stands for black or white, ``-c'' for the opposite color. Each point of the board is, of course, either black, white or empty.

A c chain is a non-empty maximum connected set of c points.

A c region is a non-empty maximum connected set of -c and empty points.

A c chain is unconditionally alive if and only if it belongs to a set of c chains S such that every chain B in S is adjacent to at least two c regions R that satisfy:

\begin{enumerate}
  \item (P1) all empty points in R are adjacent to B, and

  \item (P2) all c chains adjacent to R belong to S.
\end{enumerate}

\chapter{Engine choice}
There are a number of Go engines, so we sought to extend an existing engine rather than write something completely new. We could not find an engine in JavaScript, so the first order of business was to choose an engine to port. Our criteria for choosing an engine was simple: What is the engine with the smallest (estimated) expression in JavaScript and the highest kyu ranking?

SimpleGo v1

50kyu greedy search (random move from list of moves with worst opponent score)

SimpleGo v3

30kyu $\alpha$-$\beta$ pruned search

Crawler

?? (~30kyu) maintain line strength/shape

\chapter{Technical details}
Code Listing
- Implementation of Benson's algorithm for determining unconditional life
- Greedy Search algorithm: select\_scored\_move
- Crawler algorithm: select\_crawler\_move

%(5)    Discussions, observations, and comparisons.
%(6)    Recommendations, especially for future work and unsolved issues.
%(7)    References.

\bibliography{paper}
\bibliographystyle{plain}
\end{document}

### 2007-12-19 simple_go1/jaigo.compressed.js
added shrinksafe'd combined js file.

@@ -1,0 +1,1 @@
+function debug_output(_1,_2,_3){if(arguments.length<2){_2="debug_output";}var _4=document.getElementById(_2);if(arguments.length<3){_4.value+=_1+"\n";}else{_4.value+="<"+_3+">"+_1+"</"+_3+">";}}String.prototype.strip=(function(){var _5=/(\[|\\)/g;var ws=/(^\s*)(.*?)(\s*$)/;var _7=function(_8){var _9;if(!_8){_9=ws;}else{var _a="["+_8.replace(_5,"\\$&")+"]";_9=new RegExp("(^"+_a+"*)(.*?)("+_a+"*$)");}return this.replace(_9,"$2");};return _7;})();function diagram2game(_b){var _c=/^\d*$/;var g;var _e=_b.split("\n");for(var l in _e){if(_e.hasOwnProperty(l)){var _10=/(^\s*)(.*?)(\s*$)/;var _11=_e[l].strip();if(!_11){continue;}if(g===undefined){g=new Game(_11.length);}else{if(_c.test(_11[0])){var _12=_11.split("|");var y=Number(_12[0]);var _14=_12[1];var _15=_12[2];var _16=range(_14.length);for(x in _16){if(_16.hasOwnProperty(x)){var _17=_14[x];x=Number(x)+1;if(_17===WHITE||_17===BLACK){if(g.current_board.side!==_17){g.Make_Move(PASS_MOVE);}g.Make_Move([x,y]);}}}}}}}return g;}function test_unconditional(str){var g=diagram2game(str);debug_output(g.current_board.As_String_With_Unconditional_Status());}String.prototype.Repeat=function(_1a){var s="";t=this.toString();while(--_1a>=0){s+=t;}return s;};var EMPTY=".";var BLACK="X";var WHITE="O";var WORST_SCORE=-1000000000;var normal_stone_value_ratio=0.7;var x_coords_string="ABCDEFGHJKLMNOPQRSTUVWXYZ";var max_size=x_coords_string.length;var debug_flag=false;var BOTH=BLACK+WHITE;other_side={};other_side[BLACK]=WHITE;other_side[WHITE]=BLACK;function object(o){function F(){}F.prototype=o;return new F();}deepValueEquality=function(a,b){if(arguments.length===2){a=arguments[0];b=arguments[1];}if(!(a instanceof Object)){return a===b;}for(var e in b){if(a[e]!=b[e]){return false;}}return true;};function contains(_5,_6,eq){eq=eq||deepValueEqu
[…]

### 2007-12-19 paper/paper.tex
added measurement and future work sections. expanded a little on human heuristics and concepts v computer go concepts. Added a couple of bib items, but I'm not sure if they are appropriately academic.

--- 2007-12-19 trunk/paper/paper.tex
@@
 \begin{document}
+
+\newtheorem{defn}{Definition}
 
@@
 
-The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memoize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}.
+The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}.
 The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.
+
+There are traditional concepts in go of blocks and eyes.
+Blocks are simply contiguously-connected stones of the same color.
+Eyes are empty spaces which are completely surrounded by stones of the same color.
+A good implementation of computer go will represent these concepts, as they provide vital information when generating moves, evaluating the state of the board, and in scoring the board.
 
@@
 Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
-    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function). GNUGO v Go++
+    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function).
+There is yet no standard, robust board evaluation algorithm.
+Once a main board evaluation loop is established for playing computer go and optimizations are made to the smaller evaluation functions within the loop, changing the way the board is evaluated, though potentially better for the speed, completeness, or accuracy of the evaluation, may also potentially damage the smaller evaluation functions.
     * GA \& NN seemed promising at first, but no commercial package uses them due to lack of speed.
@@
 
-A c chain is a non-empty maximum connected set of c points.
+\begin{defn}A c-chain is a non-empty maximum connected set of c points.
+\end{defn}
 
-A c region is a non-empty maximum connected set of -c and empty points.
+\begin{defn}A c-region is a non-empty maximum connected set of -c and empty points.
+\end{defn}
 
-A c chain is unconditionally alive if and only if it belongs to a set of c chains S such that every chain B in S is adjacent to at least two c regions R that satisfy:
+\begin{defn}
+A c-chain is unconditionally alive if and only if it belongs to a set of c-chains S such that every chain B in S is adjacent to at least two c-regions R that satisfy:
 
@@
 
-  \item (P2) all c chains adjacent to R belong to S.
+  \item (P2) all c-chains adjacent to R belong to S.
 \end{enumerate}
+\end{defn}
 
@@
 
-SimpleGo v1
+implemented:
 
-50kyu greedy search (random move from list of moves with worst opponent score)
+SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score.
 
-SimpleGo v3
-
-30kyu $\alpha$-$\beta$ pruned search
-
-Crawler
-
-?? (~30kyu) maintain line strength/shape
+Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.
 
@@
 Code Listing
-- Implementation of Benson's algorithm for determining unconditional life
+- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
 - Greedy Search algorithm: select\_scored\_move
@@
 
-%(5)    Discussions, observations, and comparisons.
-%(6)    Recommendations, especially for future work and unsolved issues.
+\chapter{Discussions, observations, and comparisons}
+We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities.
+Our goals were compactness and speed due to the target platform limitations.
+
+Due to issues with the jaigo board representation, our measurements were limited to the first 50 moves
+
+Regarding compactness, the total size of all jaigo files, including images, scripts, and html, was 102221 bytes for our initial release.
+The full implementation of the jaigo code is 66071 bytes \cite{SpeedReport}.
+
+Regarding speed, we do not anticipate that the choice of processor and platform of any modern PC will have a significant change to the performance of jaigo.
+
+for the first 50 moves on a 19x19 board with human (black) versus crawler (white),
+$\begin{tabular}{ccccccccc}
+\hline
+Method Name & Total Time & r1c3 & r1c4 & r1c5 & r1c6 & r1c7 & r1c8 & r1c9\\
+\hline
+(no name) & 109675 & r1c3 & r1c4 & r1c5 & r1c6 & r1c7 & r1c8 & r1c9\\
+\hline
+r2c1 & r2c2 & r2c3 & r2c4 & r2c5 & r2c6 & r2c7 & r2c8 & r2c9\\
+\hline
+r3c1 & r3c2 & r3c3 & r3c4 & r3c5 & r3c6 & r3c7 & r3c8 & r3c9\\
+\hline
+r4c1 & r4c2 & r4c3 & r4c4 & r4c5 & r4c6 & r4c7 & r4c8 & r4c9\\
+\hline
+r5c1 & r5c2 & r5c3 & r5c4 & r5c5 & r5c6 & r5c7 & r5c8 & r5c9\\
+\hline
+r6c1 & r6c2 & r6c3 & r6c4 & r6c5 & r6c6 & r6c7 & r6c8 & r6c9\\
+\hline
+r7c1 & r7c2 & r7c3 & r7c4 & r7c5 & r7c6 & r7c7 & r7c8 & r7c9\\
+\hline
+r8c1 & r8c2 & r8c3 & r8c4 & r8c5 & r8c6 & r8c7 & r8c8 & r8c9\\
+\hline
+r9c1 & r9c2 & r9c3 & r9c4 & r9c5 & r9c6 & r9c7 & r9c8 & r9c9\\
+\hline
+r10c1 & r10c2 & r10c3 & r10c4 & r10c5 & r10c6 & r10c7 & r10c8 & r10c9\\
+\hline
+r11c1 & r11c2 & r11c3 & r11c4 & r11c5 & r11c6 & r11c7 & r11c8 & r11c9\\
+\hline
+r12c1 & r12c2 & r12c3 & r12c4 & r12c5 & r12c6 & r12c7 & r12c8 & r12c9\\
+\hline
+\end{tabular}
+(no name)	109675	25.68%	14310.594ms	36908.812ms	0.337ms	0.059ms	1593.542ms	simple_go.js (line 407)
+flood_mark	1618	14.17%	7897.73ms	25481.245ms	15.749ms	0.162ms	1635.18ms	simple_go.js (line 807)
+(no name)	194678	11.75%	6551.626ms	6554.733ms	0.034ms	0.008ms	1555.243ms	simple_go.js (line 819)
+string_as_move	56065	7.48%	4168.485ms	4169.549ms	0.074ms	0.011ms	1591.844ms	simple_go.js (line 192)
+iterate	109676	7.19%	4005.749ms	22595.463ms	0.206ms	0.011ms	1593.499ms	simple_go.js (line 391)
+(no name)	283642	4.9%	2732.499ms	18589.343ms	0.066ms	0.015ms	1593.144ms	simple_go.js (line 416)
+iterateOwnProperties	5811	4.38%	2440.378ms	2847.852ms	0.49ms	0.004ms	2.917ms	simple_go.js (line 132)
+Set_Goban	2150	3.53%	1968.651ms	1969.095ms	0.916ms	0.004ms	942.058ms	simple_go.js (line 528)
+(no name)	42734	2.99%	1666.478ms	1666.716ms	0.039ms	0.008ms	935.83ms	simple_go.js (line 835)
+update	3662	2.9%	1613.61ms	1614.099ms	0.441ms	0.004ms	1560.416ms	simple_go.js (line 117)
+Add_Block	2340	2.8%	1559.92ms	1560.774ms	0.667ms	0.007ms	1529.63ms	simple_go.js (line 757)
+Calculate_Neighbour	448	2.67%	1487.337ms	12205.939ms	27.245ms	0.019ms	1579.849ms	simple_go.js (line 829)
+Simple_Same_Block	2149	1.93%	1073.573ms	8185.472ms	3.809ms	0.002ms	1556.376ms	simple_go.js (line 606)
+List_Block_Liberties	3210	1.56%	871.609ms	5057.619ms	1.576ms	0.012ms	1593.058ms	simple_go.js (line 863)$
+
+for the first 50 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
+
+(no name)	$404494	32.8%	71255.203ms	147518.241ms	0.365ms	0.059ms	1650.149ms	simple_go.js (line 407)
+flood_mark	3838	15.94%	34634.022ms	96713.52ms	25.199ms	0.206ms	1736.537ms	simple_go.js (line 807)
+iterate	404495	8.75%	19004.696ms	76254.33ms	0.189ms	0.011ms	1650.101ms	simple_go.js (line 391)
+(no name)	735190	8.72%	18941.197ms	18941.987ms	0.026ms	0.008ms	1572.648ms	simple_go.js (line 819)
+string_as_move	242765	5.86%	12724.373ms	12725.279ms	0.052ms	0.011ms	1629.186ms	simple_go.js (line 192)
+(no name)	1041519	4.7%	10209.513ms	57248.372ms	0.055ms	0.015ms	1649.869ms	simple_go.js (line 416)
+List_Block_Liberties	7706	4.02%	8727.331ms	21457.809ms	2.785ms	0.013ms	1630.607ms	simple_go.js (line 863)
+iterateOwnProperties	14752	3.68%	7985.019ms	9340.503ms	0.633ms	0.004ms	2.419ms	simple_go.js (line 132)
+update	8742	3.18%	6911.102ms	6911.93ms	0.791ms	0.004ms	1648.007ms	simple_go.js (line 117)
+Calculate_Neighbour	661	3.03%	6593.032ms	60385.478ms	91.355ms	0.115ms	1679.244ms	simple_go.js (line 829)
+(no name)	160414	2.51%	5444.827ms	5445.809ms	0.034ms	0.008ms	944.464ms	simple_go.js (line 835)
+Add_Stone	6010	1.1%	2392.214ms	191461.602ms	31.857ms	1.524ms	1785.447ms	simple_go.js (line 641)$
+
+Median move response time for 10 games into the first 50 moves was was 1 second, with the high being 3 seconds and the low being unmeasurable.
+
+On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds.
+On an iPhone over AT&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
+
+Median move times in other go engines (gnugo especially)?
+
+\chapter{Recommendations, especially for future work and unsolved issues}
+We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.
+
+The performance of $\alpha-\beta$ search pruning in javascript should be investigated. like SimpleGo v3, a 30kyu $\alpha$-$\beta$ pruned search move generator.
+
+There is much more optimization and correctness work to be done to the jaigo board and game representations.
+
+The move speed of jaigo could be measured on many diverse computing platforms.
+
+Jaigo could be also used as a "widget", a popular way of packaging small javascript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.
+
+There are many open issues being tracked by the jaigo project at this time. Addressing them may lead to better, more correct performance. We anticipate continuing work on this project.
+
 %(7)    References.

### 2007-12-19 paper/paper.tex
Fleshed out current issues. Added tables for code output.

@@
 Eyes are empty spaces which are completely surrounded by stones of the same color.
-A good implementation of computer go will represent these concepts, as they provide vital information when generating moves, evaluating the state of the board, and in scoring the board.
+A good implementation of computer Go will represent these concepts, as they provide vital information when generating moves, evaluating the state of the board, and in scoring the board.
 
@@
 
-    * Finding a good evaluation function is hard.
-The human heuristics are vague, and when/how they get applied is even more vague.
+There are several significant issues to consider when programming computer Go.
+First, it is difficult to find an appropriate evaluation function.
+The human heuristics are vague, and the parameters for their application  are even more vague.
 Zobrist divided the game into phases for the purpose of selecting evaluation functions \cite{Zobrist1970b}.
-Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
-    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function).
+On the other hand, Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
+
+Second, there is a fight between optimizing the main loop -- that is, evaluating the board -- versus optimizing the small loops (given a board evaluation, local tricks to improve evaluation function).
 There is yet no standard, robust board evaluation algorithm.
-Once a main board evaluation loop is established for playing computer go and optimizations are made to the smaller evaluation functions within the loop, changing the way the board is evaluated, though potentially better for the speed, completeness, or accuracy of the evaluation, may also potentially damage the smaller evaluation functions.
-    * GA \& NN seemed promising at first, but no commercial package uses them due to lack of speed.
-    * If all old techniques are inadequate and new techniques don't pan out, whither computer go?
+If a main board evaluation loop is established for playing computer Go and optimizations are made to the smaller evaluation functions within the loop, this would change the way the board is evaluated.
+However, though such developments would potentially better the speed, completeness, or accuracy of the evaluation, they may also potentially damage the smaller evaluation functions.
 
+Third, while GA \& NN initially seemed promising, no commercial package uses them due to lack of speed.
+
+Thus, if all older techniques are inadequate and new techniques do not pan out, whither computer Go?
 Our project seeks to create a JavaScript-based Go-engine.
@@
 
-implemented:
-
+We chose to implement two different move choice engines, each of which provides a different playing experience.
 SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score.
-
 Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.
@@
 
-Due to issues with the jaigo board representation, our measurements were limited to the first 50 moves
+Due to issues with the jaigo board representation, our measurements were limited to the first 100 moves.
 
@@
 
-for the first 50 moves on a 19x19 board with human (black) versus crawler (white),
+For the first 50 moves on a 19x19 board with human (black) versus crawler (white),
 $\begin{tabular}{ccccccccc}
 \hline
-Method Name & Total Time & r1c3 & r1c4 & r1c5 & r1c6 & r1c7 & r1c8 & r1c9\\
+Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\
 \hline
-(no name) & 109675 & r1c3 & r1c4 & r1c5 & r1c6 & r1c7 & r1c8 & r1c9\\
+(no name) & 109675 & 25.68% & 14310.594ms &	36908.812ms &	0.337ms &	0.059ms	& 1593.542ms	& simple_go.js (line 407)\\
 \hline
-r2c1 & r2c2 & r2c3 & r2c4 & r2c5 & r2c6 & r2c7 & r2c8 & r2c9\\
+flood_mark & 1618 & 14.17% & 7897.73ms & 25481.245ms & 15.749ms & 0.162ms & 1635.18ms & simple_go.js (line 807)\\
 \hline
-r3c1 & r3c2 & r3c3 & r3c4 & r3c5 & r3c6 & r3c7 & r3c8 & r3c9\\
+(no name) &	194678 &	11.75%	& 6551.626ms &	6554.733ms &	0.034ms &	0.008ms &	1555.243ms &	simple_go.js (line 819)\\
 \hline
-r4c1 & r4c2 & r4c3 & r4c4 & r4c5 & r4c6 & r4c7 & r4c8 & r4c9\\
+string_as_move &	56065 &	7.48% &	4168.485ms &	4169.549ms &	0.074ms &	0.011ms &	1591.844ms &	simple_go.js (line 192)\\
 \hline
-r5c1 & r5c2 & r5c3 & r5c4 & r5c5 & r5c6 & r5c7 & r5c8 & r5c9\\
+iterate	109676 &	7.19% &	4005.749ms &	22595.463ms &	0.206ms &	0.011ms &	1593.499ms &	simple_go.js (line 391)\\
 \hline
-r6c1 & r6c2 & r6c3 & r6c4 & r6c5 & r6c6 & r6c7 & r6c8 & r6c9\\
+(no name) &	283642 &	4.9% &	2732.499ms &	18589.343ms &	0.066ms &	0.015ms &	1593.144ms &	simple_go.js & (line 416)\\
 \hline
-r7c1 & r7c2 & r7c3 & r7c4 & r7c5 & r7c6 & r7c7 & r7c8 & r7c9\\
+iterateOwnProperties &	5811	& 4.38%	& 2440.378ms &	2847.852ms &	0.49ms &	0.004ms &	2.917ms &	simple_go.js (line 132)\\
 \hline
-r8c1 & r8c2 & r8c3 & r8c4 & r8c5 & r8c6 & r8c7 & r8c8 & r8c9\\
+Set_Goban &	2150 &	3.53% &	1968.651ms &	1969.095ms &	0.916ms &	0.004ms &	942.058ms &	simple_go.js (line 528)\\
 \hline
-r9c1 & r9c2 & r9c3 & r9c4 & r9c5 & r9c6 & r9c7 & r9c8 & r9c9\\
+(no name) &	42734 &	2.99% &	1666.478ms &	1666.716ms &	0.039ms &	0.008ms &	935.83ms &	simple_go.js (line 835)\\
 \hline
-r10c1 & r10c2 & r10c3 & r10c4 & r10c5 & r10c6 & r10c7 & r10c8 & r10c9\\
+update &	3662 &	2.9% &	1613.61ms &	1614.099ms &	0.441ms &	0.004ms &	1560.416ms &	simple_go.js (line 117)\\
 \hline
-r11c1 & r11c2 & r11c3 & r11c4 & r11c5 & r11c6 & r11c7 & r11c8 & r11c9\\
+Add_Block &	2340 &	2.8% &	1559.92ms &	1560.774ms &	0.667ms &	0.007ms &	1529.63ms &	simple_go.js (line 757)\\
 \hline
-r12c1 & r12c2 & r12c3 & r12c4 & r12c5 & r12c6 & r12c7 & r12c8 & r12c9\\
+Calculate_Neighbour &	448	2.67% &	1487.337ms &	12205.939ms &	27.245ms &	0.019ms &	1579.849ms &	simple_go.js (line 829)\\
+\hline
+Simple_Same_Block &	2149 &	1.93% &	1073.573ms &	8185.472ms &	3.809ms &	0.002ms &	1556.376ms &	simple_go.js (line 606)\\
+\hline
+List_Block_Liberties &	3210 &	1.56% &	871.609ms &	5057.619ms &	1.576ms &	0.012ms &	1593.058ms &	simple_go.js (line 863)\\
 \hline
 \end{tabular}
-(no name)	109675	25.68%	14310.594ms	36908.812ms	0.337ms	0.059ms	1593.542ms	simple_go.js (line 407)
-flood_mark	1618	14.17%	7897.73ms	25481.245ms	15.749ms	0.162ms	1635.18ms	simple_go.js (line 807)
-(no name)	194678	11.75%	6551.626ms	6554.733ms	0.034ms	0.008ms	1555.243ms	simple_go.js (line 819)
-string_as_move	56065	7.48%	4168.485ms	4169.549ms	0.074ms	0.011ms	1591.844ms	simple_go.js (line 192)
-iterate	109676	7.19%	4005.749ms	22595.463ms	0.206ms	0.011ms	1593.499ms	simple_go.js (line 391)
-(no name)	283642	4.9%	2732.499ms	18589.343ms	0.066ms	0.015ms	1593.144ms	simple_go.js (line 416)
-iterateOwnProperties	5811	4.38%	2440.378ms	2847.852ms	0.49ms	0.004ms	2.917ms	simple_go.js (line 132)
-Set_Goban	2150	3.53%	1968.651ms	1969.095ms	0.916ms	0.004ms	942.058ms	simple_go.js (line 528)
-(no name)	42734	2.99%	1666.478ms	1666.716ms	0.039ms	0.008ms	935.83ms	simple_go.js (line 835)
-update	3662	2.9%	1613.61ms	1614.099ms	0.441ms	0.004ms	1560.416ms	simple_go.js (line 117)
-Add_Block	2340	2.8%	1559.92ms	1560.774ms	0.667ms	0.007ms	1529.63ms	simple_go.js (line 757)
-Calculate_Neighbour	448	2.67%	1487.337ms	12205.939ms	27.245ms	0.019ms	1579.849ms	simple_go.js (line 829)
-Simple_Same_Block	2149	1.93%	1073.573ms	8185.472ms	3.809ms	0.002ms	1556.376ms	simple_go.js (line 606)
-List_Block_Liberties	3210	1.56%	871.609ms	5057.619ms	1.576ms	0.012ms	1593.058ms	simple_go.js (line 863)$
 
-for the first 50 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
+For the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
 
-(no name)	$404494	32.8%	71255.203ms	147518.241ms	0.365ms	0.059ms	1650.149ms	simple_go.js (line 407)
-flood_mark	3838	15.94%	34634.022ms	96713.52ms	25.199ms	0.206ms	1736.537ms	simple_go.js (line 807)
-iterate	404495	8.75%	19004.696ms	76254.33ms	0.189ms	0.011ms	1650.101ms	simple_go.js (line 391)
-(no name)	735190	8.72%	18941.197ms	18941.987ms	0.026ms	0.008ms	1572.648ms	simple_go.js (line 819)
-string_as_move	242765	5.86%	12724.373ms	12725.279ms	0.052ms	0.011ms	1629.186ms	simple_go.js (line 192)
-(no name)	1041519	4.7%	10209.513ms	57248.372ms	0.055ms	0.015ms	1649.869ms	simple_go.js (line 416)
-List_Block_Liberties	7706	4.02%	8727.331ms	21457.809ms	2.785ms	0.013ms	1630.607ms	simple_go.js (line 863)
-iterateOwnProperties	14752	3.68%	7985.019ms	9340.503ms	0.633ms	0.004ms	2.419ms	simple_go.js (line 132)
-update	8742	3.18%	6911.102ms	6911.93ms	0.791ms	0.004ms	1648.007ms	simple_go.js (line 117)
-Calculate_Neighbour	661	3.03%	6593.032ms	60385.478ms	91.355ms	0.115ms	1679.244ms	simple_go.js (line 829)
-(no name)	160414	2.51%	5444.827ms	5445.809ms	0.034ms	0.008ms	944.464ms	simple_go.js (line 835)
-Add_Stone	6010	1.1%	2392.214ms	191461.602ms	31.857ms	1.524ms	1785.447ms	simple_go.js (line 641)$
+\begin{tabular}{ccccccccc}
+	\hline
+	Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	
+\hline
+(no name) &	$404494 &	32.8% &	71255.203ms &	147518.241ms &	0.365ms &	0.059ms &	1650.149ms &	simple_go.js (line 407)\\
+\hline
+flood_mark &	3838 &	15.94% &	34634.022ms &	96713.52ms &	25.199ms &	0.206ms &	1736.537ms &	simple_go.js (line 807)\\
+\hline
+iterate &	404495 &	8.75% &	19004.696ms &	76254.33ms &	0.189ms &	0.011ms &	1650.101ms &	simple_go.js (line 391)\\
+\hline
+(no name) &	735190 &	8.72% &	18941.197ms &	18941.987ms &	0.026ms &	0.008ms &	1572.648ms &	simple_go.js (line 819)\\
+\hline
+string_as_move &	242765 &	5.86% &	12724.373ms	& 12725.279ms &	0.052ms &	0.011ms &	1629.186ms &	simple_go.js (line 192)\\
+\hline
+(no name) &	1041519 &	4.7% &	10209.513ms &	57248.372ms &	0.055ms &	0.015ms &	1649.869ms &	simple_go.js (line 416)\\
+\hline
+List_Block_Liberties &	7706 &	4.02% &	8727.331ms &	21457.809ms &	2.785ms &	0.013ms &	1630.607ms &	simple_go.js (line 863)\\
+\hline
+iterateOwnProperties &	14752 &	3.68% &	7985.019ms &	9340.503ms &	0.633ms &	0.004ms &	2.419ms &	simple_go.js (line 132)\\
+\hline
+update &	8742 &	3.18% &	6911.102ms &	6911.93ms &	0.791ms &	0.004ms &	1648.007ms &	simple_go.js (line 117)\\
+\hline
+Calculate_Neighbour &	661 &	3.03% &	6593.032ms &	60385.478ms &	91.355ms &	0.115ms &	1679.244ms &	simple_go.js (line 829)\\
+\hline
+(no name) &	160414 &	2.51% &	5444.827ms &	5445.809ms &	0.034ms	& 0.008ms &	944.464ms &	simple_go.js (line 835)\\
+\hline
+Add_Stone &	6010 &	1.1% &	2392.214ms	& 191461.602ms	& 31.857ms &	1.524ms &	1785.447ms &	simple_go.js (line 641)\\
+\hline
+\end{tabular}
 
@@
 There is much more optimization and correctness work to be done to the jaigo board and game representations.
-
-The move speed of jaigo could be measured on many diverse computing platforms.
-
+For example, the move speed of jaigo could be measured on many diverse computing platforms.
 Jaigo could be also used as a "widget", a popular way of packaging small javascript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.

### 2007-12-19 trunk/paper/paper.tex
fix up LaTeX and don't version build output

@@
 
-The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memoize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}.
+The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}.
 The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.
+
+There are traditional concepts in go of blocks and eyes.
+Blocks are simply contiguously-connected stones of the same color.
+Eyes are empty spaces which are completely surrounded by stones of the same color.
+A good implementation of computer Go will represent these concepts, as they provide vital information when generating moves, evaluating the state of the board, and in scoring the board.
 
@@
 
-    * Finding a good evaluation function is hard.
-The human heuristics are vague, and when/how they get applied is even more vague.
+There are several significant issues to consider when programming computer Go.
+First, it is difficult to find an appropriate evaluation function.
+The human heuristics are vague, and the parameters for their application  are even more vague.
 Zobrist divided the game into phases for the purpose of selecting evaluation functions \cite{Zobrist1970b}.
-Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
-    * fight between optimizing the main loop (evaluating the board) vs optimizing the small loops (given a board evaluation, local tricks to improve evaluation function). GNUGO v Go++
-    * GA \& NN seemed promising at first, but no commercial package uses them due to lack of speed.
-    * If all old techniques are inadequate and new techniques don't pan out, whither computer go?
+On the other hand, Ryder distinguished between tactical and strategic utility \cite{Burmeister}.
 
+Second, there is a fight between optimizing the main loop -- that is, evaluating the board -- versus optimizing the small loops (given a board evaluation, local tricks to improve evaluation function).
+There is yet no standard, robust board evaluation algorithm.
+If a main board evaluation loop is established for playing computer Go and optimizations are made to the smaller evaluation functions within the loop, this would change the way the board is evaluated.
+However, though such developments would potentially better the speed, completeness, or accuracy of the evaluation, they may also potentially damage the smaller evaluation functions.
+
+Third, while GA \& NN initially seemed promising, no commercial package uses them due to lack of speed.
+
+Thus, if all older techniques are inadequate and new techniques do not pan out, whither computer Go?
 Our project seeks to create a JavaScript-based Go-engine.
@@
 
-SimpleGo v1
-
-50kyu greedy search (random move from list of moves with worst opponent score)
-
-SimpleGo v3
-
-30kyu $\alpha$-$\beta$ pruned search
-
-Crawler
-
-?? (~30kyu) maintain line strength/shape
+We chose to implement two different move choice engines, each of which provides a different playing experience.
+SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score.
+Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.
 
@@
 
-%(5)    Discussions, observations, and comparisons.
-%(6)    Recommendations, especially for future work and unsolved issues.
+\chapter{Discussions, observations, and comparisons}
+We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities.
+Our goals were compactness and speed due to the target platform limitations.
+
+Due to issues with the jaigo board representation, our measurements were limited to the first 100 moves.
+
+Regarding compactness, the total size of all jaigo files, including images, scripts, and html, was 102221 bytes for our initial release.
+The full implementation of the jaigo code is 66071 bytes \cite{SpeedReport}.
+
+Regarding speed, we do not anticipate that the choice of processor and platform of any modern PC will have a significant change to the performance of jaigo.
+
+For the first 50 moves on a 19x19 board with human (black) versus crawler (white),
+\begin{tabular}{ccccccccc}
+\hline
+Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\
+\hline
+(no name) & 109675 & 25.68 & 14310.594ms &	36908.812ms &	0.337ms &	0.059ms	& 1593.542ms	& simple\_go.js (line 407)\\
+\hline
+flood\_mark & 1618 & 14.17 & 7897.73ms & 25481.245ms & 15.749ms & 0.162ms & 1635.18ms & simple\_go.js (line 807)\\
+\hline
+(no name) &	194678 &	11.75	& 6551.626ms &	6554.733ms &	0.034ms &	0.008ms &	1555.243ms &	simple\_go.js (line 819)\\
+\hline
+string\_as\_move &	56065 &	7.48 &	4168.485ms &	4169.549ms &	0.074ms &	0.011ms &	1591.844ms &	simple\_go.js (line 192)\\
+\hline
+iterate	& 109676 &	7.19 &	4005.749ms &	22595.463ms &	0.206ms &	0.011ms &	1593.499ms &	simple\_go.js (line 391)\\
+\hline
+(no name) &	283642 &	4.9 &	2732.499ms &	18589.343ms &	0.066ms &	0.015ms &	1593.144ms &	simple\_go.js (line 416)\\
+\hline
+iterateOwnProperties &	5811	& 4.38	& 2440.378ms &	2847.852ms &	0.49ms &	0.004ms &	2.917ms &	simple\_go.js (line 132)\\
+\hline
+Set\_Goban &	2150 &	3.53 &	1968.651ms &	1969.095ms &	0.916ms &	0.004ms &	942.058ms &	simple\_go.js (line 528)\\
+\hline
+(no name) &	42734 &	2.99 &	1666.478ms &	1666.716ms &	0.039ms &	0.008ms &	935.83ms &	simple\_go.js (line 835)\\
+\hline
+update &	3662 &	2.9 &	1613.61ms &	1614.099ms &	0.441ms &	0.004ms &	1560.416ms &	simple\_go.js (line 117)\\
+\hline
+Add\_Block &	2340 &	2.8 &	1559.92ms &	1560.774ms &	0.667ms &	0.007ms &	1529.63ms &	simple\_go.js (line 757)\\
+\hline
+Calculate\_Neighbour &	448	2.67 &	1487.337ms &	12205.939ms &	27.245ms &	0.019ms &	1579.849ms &	simple\_go.js (line 829)\\
+\hline
+Simple\_Same\_Block &	2149 &	1.93 &	1073.573ms &	8185.472ms &	3.809ms &	0.002ms &	1556.376ms &	simple\_go.js (line 606)\\
+\hline
+List\_Block\_Liberties &	3210 &	1.56 &	871.609ms &	5057.619ms &	1.576ms &	0.012ms &	1593.058ms &	simple\_go.js (line 863)\\
+\hline
+\end{tabular}
+
+For the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
+
+\begin{tabular}{ccccccccc}
+	\hline
+	Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	
+\hline
+(no name) &	404494 &	32.8 &	71255.203ms &	147518.241ms &	0.365ms &	0.059ms &	1650.149ms &	simple\_go.js (line 407)\\
+\hline
+flood\_mark &	3838 &	15.94 &	34634.022ms &	96713.52ms &	25.199ms &	0.206ms &	1736.537ms &	simple\_go.js (line 807)\\
+\hline
+iterate &	404495 &	8.75 &	19004.696ms &	76254.33ms &	0.189ms &	0.011ms &	1650.101ms &	simple\_go.js (line 391)\\
+\hline
+(no name) &	735190 &	8.72 &	18941.197ms &	18941.987ms &	0.026ms &	0.008ms &	1572.648ms &	simple\_go.js (line 819)\\
+\hline
+string\_as\_move &	242765 &	5.86 &	12724.373ms	& 12725.279ms &	0.052ms &	0.011ms &	1629.186ms &	simple\_go.js (line 192)\\
+\hline
+(no name) &	1041519 &	4.7 &	10209.513ms &	57248.372ms &	0.055ms &	0.015ms &	1649.869ms &	simple\_go.js (line 416)\\
+\hline
+List\_Block\_Liberties &	7706 &	4.02 &	8727.331ms &	21457.809ms &	2.785ms &	0.013ms &	1630.607ms &	simple\_go.js (line 863)\\
+\hline
+iterateOwnProperties &	14752 &	3.68 &	7985.019ms &	9340.503ms &	0.633ms &	0.004ms &	2.419ms &	simple\_go.js (line 132)\\
+\hline
+update &	8742 &	3.18 &	6911.102ms &	6911.93ms &	0.791ms &	0.004ms &	1648.007ms &	simple\_go.js (line 117)\\
+\hline
+Calculate\_Neighbour &	661 &	3.03 &	6593.032ms &	60385.478ms &	91.355ms &	0.115ms &	1679.244ms &	simple\_go.js (line 829)\\
+\hline
+(no name) &	160414 &	2.51 &	5444.827ms &	5445.809ms &	0.034ms	& 0.008ms &	944.464ms &	simple\_go.js (line 835)\\
+\hline
+Add\_Stone &	6010 &	1.1 &	2392.214ms	& 191461.602ms	& 31.857ms &	1.524ms &	1785.447ms &	simple\_go.js (line 641)\\
+\hline
+\end{tabular}
+
+Median move response time for 10 games into the first 50 moves was was 1 second, with the high being 3 seconds and the low being unmeasurable.
+
+On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds.
+On an iPhone over AT\&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
+
+Median move times in other go engines (gnugo especially)?
+
+\chapter{Recommendations, especially for future work and unsolved issues}
+We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.
+
+The performance of $\alpha-\beta$ search pruning in javascript should be investigated. like SimpleGo v3, a 30kyu $\alpha$-$\beta$ pruned search move generator.
+
+There is much more optimization and correctness work to be done to the jaigo board and game representations.
+For example, the move speed of jaigo could be measured on many diverse computing platforms.
+Jaigo could be also used as a "widget", a popular way of packaging small javascript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.
+
+There are many open issues being tracked by the jaigo project at this time. Addressing them may lead to better, more correct performance. We anticipate continuing work on this project.
+
 %(7)    References.

### 2007-12-19 trunk/paper/paper.tex
rotate tables to avoid truncation or splitting

@@
 
+\usepackage{graphicx}
 \usepackage{hyperref}
@@
 
+\frontmatter
 \begin{titlingpage}
@@
 
+\mainmatter
 \chapter{Introduction}
@@
 
-For the first 50 moves on a 19x19 board with human (black) versus crawler (white),
-\begin{tabular}{ccccccccc}
-\hline
-Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\
-\hline
-(no name) & 109675 & 25.68 & 14310.594ms &	36908.812ms &	0.337ms &	0.059ms	& 1593.542ms	& simple\_go.js (line 407)\\
-\hline
-flood\_mark & 1618 & 14.17 & 7897.73ms & 25481.245ms & 15.749ms & 0.162ms & 1635.18ms & simple\_go.js (line 807)\\
-\hline
-(no name) &	194678 &	11.75	& 6551.626ms &	6554.733ms &	0.034ms &	0.008ms &	1555.243ms &	simple\_go.js (line 819)\\
-\hline
-string\_as\_move &	56065 &	7.48 &	4168.485ms &	4169.549ms &	0.074ms &	0.011ms &	1591.844ms &	simple\_go.js (line 192)\\
-\hline
-iterate	& 109676 &	7.19 &	4005.749ms &	22595.463ms &	0.206ms &	0.011ms &	1593.499ms &	simple\_go.js (line 391)\\
-\hline
-(no name) &	283642 &	4.9 &	2732.499ms &	18589.343ms &	0.066ms &	0.015ms &	1593.144ms &	simple\_go.js (line 416)\\
-\hline
-iterateOwnProperties &	5811	& 4.38	& 2440.378ms &	2847.852ms &	0.49ms &	0.004ms &	2.917ms &	simple\_go.js (line 132)\\
-\hline
-Set\_Goban &	2150 &	3.53 &	1968.651ms &	1969.095ms &	0.916ms &	0.004ms &	942.058ms &	simple\_go.js (line 528)\\
-\hline
-(no name) &	42734 &	2.99 &	1666.478ms &	1666.716ms &	0.039ms &	0.008ms &	935.83ms &	simple\_go.js (line 835)\\
-\hline
-update &	3662 &	2.9 &	1613.61ms &	1614.099ms &	0.441ms &	0.004ms &	1560.416ms &	simple\_go.js (line 117)\\
-\hline
-Add\_Block &	2340 &	2.8 &	1559.92ms &	1560.774ms &	0.667ms &	0.007ms &	1529.63ms &	simple\_go.js (line 757)\\
-\hline
-Calculate\_Neighbour &	448	2.67 &	1487.337ms &	12205.939ms &	27.245ms &	0.019ms &	1579.849ms &	simple\_go.js (line 829)\\
-\hline
-Simple\_Same\_Block &	2149 &	1.93 &	1073.573ms &	8185.472ms &	3.809ms &	0.002ms &	1556.376ms &	simple\_go.js (line 606)\\
-\hline
-List\_Block\_Liberties &	3210 &	1.56 &	871.609ms &	5057.619ms &	1.576ms &	0.012ms &	1593.058ms &	simple\_go.js (line 863)\\
-\hline
-\end{tabular}
+\tref{table:first50} shows the first 50 moves on a 19x19 board with human (black) versus crawler (white).
+\begin{table}
+  \rotatebox{90}{
+    \begin{tabular}{ccccccccc} \toprule
+      Function & Calls & Percent & Own Time (ms) & Time (ms) & Avg (ms) & Min (ms) & Max (ms) & File\\ \midrule
+      (no name) & 109675 & 25.68 & 14310.594 &	36908.812 &	0.337 &	0.059	& 1593.542	& simple\_go.js (line 407)\\
+      flood\_mark & 1618 & 14.17 & 7897.73 & 25481.245 & 15.749 & 0.162 & 1635.18 & simple\_go.js (line 807)\\
+      (no name) &	194678 &	11.75	& 6551.626 &	6554.733 &	0.034 &	0.008 &	1555.243 &	simple\_go.js (line 819)\\
+      string\_as\_move &	56065 &	7.48 &	4168.485 &	4169.549 &	0.074 &	0.011 &	1591.844 &	simple\_go.js (line 192)\\
+      iterate	& 109676 &	7.19 &	4005.749 &	22595.463 &	0.206 &	0.011 &	1593.499 &	simple\_go.js (line 391)\\
+      (no name) &	283642 &	4.9 &	2732.499 &	18589.343 &	0.066 &	0.015 &	1593.144 &	simple\_go.js (line 416)\\
+      iterateOwnProperties &	5811	& 4.38	& 2440.378 &	2847.852 &	0.49 &	0.004 &	2.917 &	simple\_go.js (line 132)\\
+      Set\_Goban &	2150 &	3.53 &	1968.651 &	1969.095 &	0.916 &	0.004 &	942.058 &	simple\_go.js (line 528)\\
+      (no name) &	42734 &	2.99 &	1666.478 &	1666.716 &	0.039 &	0.008 &	935.83 &	simple\_go.js (line 835)\\
+      update &	3662 &	2.9 &	1613.61 &	1614.099 &	0.441 &	0.004 &	1560.416 &	simple\_go.js (line 117)\\
+      Add\_Block &	2340 &	2.8 &	1559.92 &	1560.774 &	0.667 &	0.007 &	1529.63 &	simple\_go.js (line 757)\\
+      Calculate\_Neighbour &	448 &	2.67 &	1487.337 &	12205.939 &	27.245 &	0.019 &	1579.849 &	simple\_go.js (line 829)\\
+      Simple\_Same\_Block &	2149 &	1.93 &	1073.573 &	8185.472 &	3.809 &	0.002 &	1556.376 &	simple\_go.js (line 606)\\
+      List\_Block\_Liberties &	3210 &	1.56 &	871.609 &	5057.619 &	1.576 &	0.012 &	1593.058 &	simple\_go.js (line 863)\\ \bottomrule
+  \end{tabular}}
+  \caption{First 50 moves on a 19x19 board}
+  \label{table:first50}
+\end{table}
 
-For the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
+\tref{table:first100} shows the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white).
 
-\begin{tabular}{ccccccccc}
-	\hline
-	Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	
-\hline
-(no name) &	404494 &	32.8 &	71255.203ms &	147518.241ms &	0.365ms &	0.059ms &	1650.149ms &	simple\_go.js (line 407)\\
-\hline
-flood\_mark &	3838 &	15.94 &	34634.022ms &	96713.52ms &	25.199ms &	0.206ms &	1736.537ms &	simple\_go.js (line 807)\\
-\hline
-iterate &	404495 &	8.75 &	19004.696ms &	76254.33ms &	0.189ms &	0.011ms &	1650.101ms &	simple\_go.js (line 391)\\
-\hline
-(no name) &	735190 &	8.72 &	18941.197ms &	18941.987ms &	0.026ms &	0.008ms &	1572.648ms &	simple\_go.js (line 819)\\
-\hline
-string\_as\_move &	242765 &	5.86 &	12724.373ms	& 12725.279ms &	0.052ms &	0.011ms &	1629.186ms &	simple\_go.js (line 192)\\
-\hline
-(no name) &	1041519 &	4.7 &	10209.513ms &	57248.372ms &	0.055ms &	0.015ms &	1649.869ms &	simple\_go.js (line 416)\\
-\hline
-List\_Block\_Liberties &	7706 &	4.02 &	8727.331ms &	21457.809ms &	2.785ms &	0.013ms &	1630.607ms &	simple\_go.js (line 863)\\
-\hline
-iterateOwnProperties &	14752 &	3.68 &	7985.019ms &	9340.503ms &	0.633ms &	0.004ms &	2.419ms &	simple\_go.js (line 132)\\
-\hline
-update &	8742 &	3.18 &	6911.102ms &	6911.93ms &	0.791ms &	0.004ms &	1648.007ms &	simple\_go.js (line 117)\\
-\hline
-Calculate\_Neighbour &	661 &	3.03 &	6593.032ms &	60385.478ms &	91.355ms &	0.115ms &	1679.244ms &	simple\_go.js (line 829)\\
-\hline
-(no name) &	160414 &	2.51 &	5444.827ms &	5445.809ms &	0.034ms	& 0.008ms &	944.464ms &	simple\_go.js (line 835)\\
-\hline
-Add\_Stone &	6010 &	1.1 &	2392.214ms	& 191461.602ms	& 31.857ms &	1.524ms &	1785.447ms &	simple\_go.js (line 641)\\
-\hline
-\end{tabular}
+\begin{table}
+  \rotatebox{90}{
+    \begin{tabular}{ccccccccc} \toprule
+      Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	\midrule
+      (no name) &	404494 &	32.8 &	71255.203 &	147518.241 &	0.365 &	0.059 &	1650.149 &	simple\_go.js (line 407)\\
+      flood\_mark &	3838 &	15.94 &	34634.022 &	96713.52 &	25.199 &	0.206 &	1736.537 &	simple\_go.js (line 807)\\
+      iterate &	404495 &	8.75 &	19004.696 &	76254.33 &	0.189 &	0.011 &	1650.101 &	simple\_go.js (line 391)\\
+      (no name) &	735190 &	8.72 &	18941.197 &	18941.987 &	0.026 &	0.008 &	1572.648 &	simple\_go.js (line 819)\\
+      string\_as\_move &	242765 &	5.86 &	12724.373	& 12725.279 &	0.052 &	0.011 &	1629.186 &	simple\_go.js (line 192)\\
+      (no name) &	1041519 &	4.7 &	10209.513 &	57248.372 &	0.055 &	0.015 &	1649.869 &	simple\_go.js (line 416)\\
+      List\_Block\_Liberties &	7706 &	4.02 &	8727.331 &	21457.809 &	2.785 &	0.013 &	1630.607 &	simple\_go.js (line 863)\\
+      iterateOwnProperties &	14752 &	3.68 &	7985.019 &	9340.503 &	0.633 &	0.004 &	2.419 &	simple\_go.js (line 132)\\
+      update &	8742 &	3.18 &	6911.102 &	6911.93 &	0.791 &	0.004 &	1648.007 &	simple\_go.js (line 117)\\
+      Calculate\_Neighbour &	661 &	3.03 &	6593.032 &	60385.478 &	91.355 &	0.115 &	1679.244 &	simple\_go.js (line 829)\\
+      (no name) &	160414 &	2.51 &	5444.827 &	5445.809 &	0.034	& 0.008 &	944.464 &	simple\_go.js (line 835)\\
+      Add\_Stone &	6010 &	1.1 &	2392.214	& 191461.602	& 31.857 &	1.524 &	1785.447 &	simple\_go.js (line 641)\\ \bottomrule
+    \end{tabular}}
+  \caption{First 100 moves on a 19x19 board}
+  \label{table:first100}
+\end{table}
 
@@
 
-%(7)    References.
-
+\backmatter
 \bibliography{paper}

### 2007-12-19 trunk/paper/paper.tex
editing for style and punctuation

@@
 
-While chess eventually yielded to clever hardware and algorithms, the sheer state-space of Go has rendered specialize search approaches weak. In 1965, noted statistician I.J. Good noted presciently:
+While chess eventually yielded to clever hardware and algorithms, the sheer state-space of Go has rendered specialize search approaches weak.
+In 1965, noted statistician I.J.
+Good presciently commented:
 
@@
 Go on a computer?
-– In order to programme [sic] a computer to play a reasonable game of Go – rather than merely a legal game – it is necessary to formalise the principles of good strategy, or to design a learning programme.
+-- In order to programme [sic] a computer to play a reasonable game of Go -- rather than merely a legal game -- it is necessary to formalise the principles of good strategy, or to design a learning programme.
 The principles are more qualitative and mysterious than in chess, and depend more on judgment.
@@
 The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}.
-The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.
+The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), $f: (x, y, color) \rightarrow Z$.
 
@@
 
-Third, while GA \& NN initially seemed promising, no commercial package uses them due to lack of speed.
+Third, while genetic algorithms and neural networks initially seemed promising, no commercial package uses them due to lack of speed.
 
@@
 It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail.
-As professional programmers, we viewed this project as an opportunity to explore not only AI but also renewed interest in the JavaScript language.
+As professional programmers, we viewed this project as an opportunity to explore not only AI but also to pursue our renewed interest in the JavaScript language.
 Also, JavaScript is the only supported means of developing iPhone applications.
-iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations only heighten the interest.
+iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations heighten the interest.
 
@@
 
-We chose to implement two different move choice engines, each of which provides a different playing experience.
+We implemented two different move choice engines, each of which provides a different playing experience.
 SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score.
@@
 \chapter{Technical details}
-Code Listing
+%Code Listing
 - Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
@@
 \chapter{Discussions, observations, and comparisons}
-We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities.
+We implemented representations of the board, the game, eyes, and blocks; scoring for color and blocks based on Benson's algorithm; two move choice engines based on two different algorithms; and numerous language-specific utilities.
 Our goals were compactness and speed due to the target platform limitations.
@@
 \chapter{Recommendations, especially for future work and unsolved issues}
-We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.
+We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo.
+Our anecdotal experience and those of others on the Web show that there is an order of magnitude difference between popular browsers for some JavaScript operations.
+However, we have not carefully studied this issue.%http://therealcrisp.xs4all.nl/blog/2006/12/09/string-performance-in-internet-explorer/
 

### 2007-12-19 paper/crawler.js
incorporate selected listings

@@ -1,0 +1,56 @@
+
+Game.prototype.Select_Crawler_Move = function(remove_opponent_dead, pass_allowed)
+{
+	if (arguments.length < 2) 
+        {
+                if (arguments.length < 1) 
+                {
+                remove_opponent_dead = false;
+                }
+                pass_allowed = true;
+    }
+    var args = KeywordArguments.prototype.combine(arguments, {remove_opponent_dead: remove_opponent_dead, pass_allowed : pass_allowed});
+    
+	/* find the move that produces the largest liberties.
+	   reject a move if it results in less liberties for the block than before.
+	*/
+	var best_moves = [];
+	var most_liberties = 0;
+	
+	if (this.move_history.length <= 1)
+	{
+		//for the first move, do something intelligent in the middle.
+		var middle_moves = this.current_board.List_Empty_3x3_Neighbour([9,9]);
+		return randomchoice(middle_moves[0]);
+	}
+	
+	for (var b in this.current_board.block_list)
+	{
+		if (this.current_board.block_list[b].color === this.current_board.side)
+		{
+			var liberties = this.current_board.List_Block_Liberties(this.current_board.block_list[b]);
+			for (var liberty in liberties)
+			{
+				var current_block_liberties = this.current_board.Block_Liberties(this.current_board.block_list[b]);
+				if (current_block_liberties >= most_liberties)
+				{
+					this.Make_Move(liberties[liberty]);
+					var new_block_liberties = this.current_board.Block_Liberties(this.current_board.block_list[b]);
+					if (new_block_liberties >= current_block_liberties)
+					{
+						best_moves.push(liberties[liberty]);
+					}
+					this.Undo_Move();
+				}
+			}
+		}
+	}
+	if (best_moves.length > 0)
+	{
+		return randomchoice(best_moves);
+	}
+	else
+	{
+		return this.Select_Random_Move();
+	}
+};

### 2007-12-19 trunk/paper/paper.tex
incorporate selected listings

@@
 
+\usepackage{listings}
 \usepackage{graphicx}
@@
 \listoftables
+\lstlistoflistings
 
@@
 %Code Listing
-- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
-- Greedy Search algorithm: select\_scored\_move
-- Crawler algorithm: select\_crawler\_move
+\lstset{language=Java,breaklines=true}% if I have extra time, I'll define JavaScript for real
+\lstinputlisting[label={code:benson}, caption={Benson's Algorithm}]{benson.js}
+\lstinputlisting[label={code:greedy}, caption={Move generator: greedy search}]{greedy.js}
+\lstinputlisting[label={code:crawler}, caption={Move generator: crawler}]{crawler.js}
 

### 2007-12-19 trunk/paper/benson.js

@@ -1,0 +1,268 @@
+Board.prototype.Analyze_Color_Unconditional_Status = function(color)
+{
+    //find potential eyes
+    var eye_colors = EMPTY + other_side[color];
+    var eye_list = this.listPotentialEyes(eye_colors);
+    var board = this;
+    //check that all empty points are adjacent to our color
+    var lists = this.classifyPotentialEyesByColor(eye_list, color);
+    var not_ok_eye_list; //these might still contain dead groups if totally inside live group
+    eye_list = lists[0];
+    not_ok_eye_list = lists[1];
+
+    //first we assume all blocks to be ok
+    this.each_Blocks(color)(function (block) {
+            block.eye_count = 2;
+        });
+    //main loop: at end of loop check if changes
+    while (true) {
+        var changed_count = 0;
+        this.each_Blocks(color)(function (block) {
+                //not really needed but short and probably useful optimization
+                if (block.eye_count < 2) {
+                    return board.each_Blocks['continue'];
+                }
+                //count eyes
+                var block_eye_list = [];
+                for (var stone in block.neighbour)
+                {
+                        var eye = board.blocks[stone].eye;
+                        if (eye && block_eye_list.indexOf(eye) == -1)
+                    { 
+                                block_eye_list.push(eye);
+                    }
+                }
+                //count only those eyespaces which have empty point(s) adjacent to this block
+                block.eye_count = 0;
+                for (var e in block_eye_list) {
+                    var eye = block_eye_list[e];
+                    if (contains(eye.our_blocks, block, deepValueEquality)) {
+                        block.eye_count = block.eye_count + 1;
+                    }
+                }
+                if (block.eye_count < 2) {
+                    changed_count = changed_count + 1;
+                }
+            });
+        //check eyes for required all groups 2 eyes
+        var ok_eye_list = [];
+        for (var e in eye_list) {
+            var eye = eye_list[e];
+            eye_is_ok = true;
+            this.each_Neighbour_Eye_Blocks(eye)(function (block) {
+                    if (block.eye_count < 2) {
+                        eye_is_ok = false;
+                        return board.each_Neighbour_Eye_Blocks['break'];
+                    }
+                });
+            if (eye_is_ok) {
+                ok_eye_list.push(eye);
+            }
+            else {
+                changed_count = changed_count + 1;
+                not_ok_eye_list.push(eye);
+                //remove reference to eye that is not ok
+                for (var b in eye.parts) {
[…]

### 2007-12-19 trunk/paper/greedy.js

@@ -1,0 +1,82 @@
+Game.prototype.Select_Scored_Move = function(remove_opponent_dead, pass_allowed) {
+    if (arguments.length < 2) 
+        {
+                if (arguments.length < 1) 
+                {
+                remove_opponent_dead = false;
+                }
+                pass_allowed = true;
+    }
+    var args = KeywordArguments.prototype.combine(arguments, {remove_opponent_dead: remove_opponent_dead, pass_allowed : pass_allowed});
+
+    /* Go through all legal moves.
+       Keep track of best score and all moves that achieve it.
+       Select one move from best moves and return it.
+    */
+    var territory_moves_forbidden = args['pass_allowed'];
+    var base_score = this.current_board.Score_Position();
+    if (debug_flag)
+        {
+        //print "?", base_score
+        }
+    //if abs(base_score)==self.size**2:
+    //    import pdb; pdb.set_trace()
+    var has_unknown_status_block = this.current_board.Has_Block_Status(WHITE+BLACK+EMPTY, "unknown");
+    var has_opponent_dead_block = this.current_board.Has_Block_Status(other_side[this.current_board.side], "dead");
+    //has unsettled blocks
+        if (has_unknown_status_block) {
+            args['pass_allowed'] = false;
+        }
+    //dead removal has been requested and there are dead opponent stones
+    if (args['remove_opponent_dead'] && has_opponent_dead_block)
+        {
+        territory_moves_forbidden = false;
+        args['pass_allowed'] = false;
+        }
+    var forbidden_moves =
+        territory_moves_forbidden ?
+        this.current_board.Territory_As_Dict() :
+        [];
+    var best_score = WORST_SCORE;
+    var best_moves = [];
+        this.each_Moves()(function (move) {
+                if (forbidden_moves[move] !== undefined) {
+                    return this.each_Moves['continue'];
+                }
+                score = -Game.prototype.Score_Move(move) - base_score;
+                //self.make_move(move)
+                //get score from our viewpoint: negative of opponent score
+                //score = -self.current_board.score_position() - base_score
+                //score = -self.score_position() - base_score
+                if (debug_flag)
+                    {
+                        debug_output(score + move_as_string(move, this.size));
+                    }
+                //self.undo_move()
+                //Give pass move slight bonus so its preferred among do nothing moves
+                if (deepValueEquality(move, PASS_MOVE))
+                    {
+                        if (!args['pass_allowed']) {
+                            return this.each_Moves['continue'];
+                        }
+                        score = score + 0.001;
+                    }
[…]

### 2007-12-19 paper/paper.tex
todo

@@
 
+\usepackage{listings}
+\usepackage{graphicx}
 \usepackage{hyperref}
@@
 \listoftables
+\lstlistoflistings
 
+\mainmatter
 \chapter{Introduction}
@@
 \chapter{Technical details}
-Code Listing
-- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
-- Greedy Search algorithm: select\_scored\_move
-- Crawler algorithm: select\_crawler\_move
+%Code Listing
+\lstset{language=Java,breaklines=true}% if I have extra time, I'll define JavaScript for real
+\lstinputlisting[label={code:benson}, caption={Benson's Algorithm}]{benson.js}
+\lstinputlisting[label={code:greedy}, caption={Move generator: greedy search}]{greedy.js}
+\lstinputlisting[label={code:crawler}, caption={Move generator: crawler}]{crawler.js}
 
 \chapter{Discussions, observations, and comparisons}
-We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities.
+We implemented representations of the board, the game, eyes, and blocks; scoring for color and blocks based on Benson's algorithm; two move choice engines based on two different algorithms; and numerous language-specific utilities.
 Our goals were compactness and speed due to the target platform limitations.
@@
 
-For the first 50 moves on a 19x19 board with human (black) versus crawler (white),
-$\begin{tabular}{ccccccccc}
-\hline
-Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\
-\hline
-(no name) & 109675 & 25.68% & 14310.594ms &	36908.812ms &	0.337ms &	0.059ms	& 1593.542ms	& simple_go.js (line 407)\\
-\hline
-flood_mark & 1618 & 14.17% & 7897.73ms & 25481.245ms & 15.749ms & 0.162ms & 1635.18ms & simple_go.js (line 807)\\
-\hline
-(no name) &	194678 &	11.75%	& 6551.626ms &	6554.733ms &	0.034ms &	0.008ms &	1555.243ms &	simple_go.js (line 819)\\
-\hline
-string_as_move &	56065 &	7.48% &	4168.485ms &	4169.549ms &	0.074ms &	0.011ms &	1591.844ms &	simple_go.js (line 192)\\
-\hline
-iterate	109676 &	7.19% &	4005.749ms &	22595.463ms &	0.206ms &	0.011ms &	1593.499ms &	simple_go.js (line 391)\\
-\hline
-(no name) &	283642 &	4.9% &	2732.499ms &	18589.343ms &	0.066ms &	0.015ms &	1593.144ms &	simple_go.js & (line 416)\\
-\hline
-iterateOwnProperties &	5811	& 4.38%	& 2440.378ms &	2847.852ms &	0.49ms &	0.004ms &	2.917ms &	simple_go.js (line 132)\\
-\hline
-Set_Goban &	2150 &	3.53% &	1968.651ms &	1969.095ms &	0.916ms &	0.004ms &	942.058ms &	simple_go.js (line 528)\\
-\hline
-(no name) &	42734 &	2.99% &	1666.478ms &	1666.716ms &	0.039ms &	0.008ms &	935.83ms &	simple_go.js (line 835)\\
-\hline
-update &	3662 &	2.9% &	1613.61ms &	1614.099ms &	0.441ms &	0.004ms &	1560.416ms &	simple_go.js (line 117)\\
-\hline
-Add_Block &	2340 &	2.8% &	1559.92ms &	1560.774ms &	0.667ms &	0.007ms &	1529.63ms &	simple_go.js (line 757)\\
-\hline
-Calculate_Neighbour &	448	2.67% &	1487.337ms &	12205.939ms &	27.245ms &	0.019ms &	1579.849ms &	simple_go.js (line 829)\\
-\hline
-Simple_Same_Block &	2149 &	1.93% &	1073.573ms &	8185.472ms &	3.809ms &	0.002ms &	1556.376ms &	simple_go.js (line 606)\\
-\hline
-List_Block_Liberties &	3210 &	1.56% &	871.609ms &	5057.619ms &	1.576ms &	0.012ms &	1593.058ms &	simple_go.js (line 863)\\
-\hline
-\end{tabular}
+\tref{table:first50} shows the first 50 moves on a 19x19 board with human (black) versus crawler (white).
+\begin{table}
+  \rotatebox{90}{
+    \begin{tabular}{ccccccccc} \toprule
+      Function & Calls & Percent & Own Time (ms) & Time (ms) & Avg (ms) & Min (ms) & Max (ms) & File\\ \midrule
+      (no name) & 109675 & 25.68 & 14310.594 &	36908.812 &	0.337 &	0.059	& 1593.542	& simple\_go.js (line 407)\\
+      flood\_mark & 1618 & 14.17 & 7897.73 & 25481.245 & 15.749 & 0.162 & 1635.18 & simple\_go.js (line 807)\\
+      (no name) &	194678 &	11.75	& 6551.626 &	6554.733 &	0.034 &	0.008 &	1555.243 &	simple\_go.js (line 819)\\
+      string\_as\_move &	56065 &	7.48 &	4168.485 &	4169.549 &	0.074 &	0.011 &	1591.844 &	simple\_go.js (line 192)\\
+      iterate	& 109676 &	7.19 &	4005.749 &	22595.463 &	0.206 &	0.011 &	1593.499 &	simple\_go.js (line 391)\\
+      (no name) &	283642 &	4.9 &	2732.499 &	18589.343 &	0.066 &	0.015 &	1593.144 &	simple\_go.js (line 416)\\
+      iterateOwnProperties &	5811	& 4.38	& 2440.378 &	2847.852 &	0.49 &	0.004 &	2.917 &	simple\_go.js (line 132)\\
+      Set\_Goban &	2150 &	3.53 &	1968.651 &	1969.095 &	0.916 &	0.004 &	942.058 &	simple\_go.js (line 528)\\
+      (no name) &	42734 &	2.99 &	1666.478 &	1666.716 &	0.039 &	0.008 &	935.83 &	simple\_go.js (line 835)\\
+      update &	3662 &	2.9 &	1613.61 &	1614.099 &	0.441 &	0.004 &	1560.416 &	simple\_go.js (line 117)\\
+      Add\_Block &	2340 &	2.8 &	1559.92 &	1560.774 &	0.667 &	0.007 &	1529.63 &	simple\_go.js (line 757)\\
+      Calculate\_Neighbour &	448 &	2.67 &	1487.337 &	12205.939 &	27.245 &	0.019 &	1579.849 &	simple\_go.js (line 829)\\
+      Simple\_Same\_Block &	2149 &	1.93 &	1073.573 &	8185.472 &	3.809 &	0.002 &	1556.376 &	simple\_go.js (line 606)\\
+      List\_Block\_Liberties &	3210 &	1.56 &	871.609 &	5057.619 &	1.576 &	0.012 &	1593.058 &	simple\_go.js (line 863)\\ \bottomrule
+  \end{tabular}}
+  \caption{First 50 moves on a 19x19 board}
+  \label{table:first50}
+\end{table}
 
-For the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white),
+\tref{table:first100} shows the first 100 moves of a game on a 19x19 board, with crawler (black) versus crawler (white).
 
-\begin{tabular}{ccccccccc}
-	\hline
-	Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	
-\hline
-(no name) &	$404494 &	32.8% &	71255.203ms &	147518.241ms &	0.365ms &	0.059ms &	1650.149ms &	simple_go.js (line 407)\\
-\hline
-flood_mark &	3838 &	15.94% &	34634.022ms &	96713.52ms &	25.199ms &	0.206ms &	1736.537ms &	simple_go.js (line 807)\\
-\hline
-iterate &	404495 &	8.75% &	19004.696ms &	76254.33ms &	0.189ms &	0.011ms &	1650.101ms &	simple_go.js (line 391)\\
-\hline
-(no name) &	735190 &	8.72% &	18941.197ms &	18941.987ms &	0.026ms &	0.008ms &	1572.648ms &	simple_go.js (line 819)\\
-\hline
-string_as_move &	242765 &	5.86% &	12724.373ms	& 12725.279ms &	0.052ms &	0.011ms &	1629.186ms &	simple_go.js (line 192)\\
-\hline
-(no name) &	1041519 &	4.7% &	10209.513ms &	57248.372ms &	0.055ms &	0.015ms &	1649.869ms &	simple_go.js (line 416)\\
-\hline
-List_Block_Liberties &	7706 &	4.02% &	8727.331ms &	21457.809ms &	2.785ms &	0.013ms &	1630.607ms &	simple_go.js (line 863)\\
-\hline
-iterateOwnProperties &	14752 &	3.68% &	7985.019ms &	9340.503ms &	0.633ms &	0.004ms &	2.419ms &	simple_go.js (line 132)\\
-\hline
-update &	8742 &	3.18% &	6911.102ms &	6911.93ms &	0.791ms &	0.004ms &	1648.007ms &	simple_go.js (line 117)\\
-\hline
-Calculate_Neighbour &	661 &	3.03% &	6593.032ms &	60385.478ms &	91.355ms &	0.115ms &	1679.244ms &	simple_go.js (line 829)\\
-\hline
-(no name) &	160414 &	2.51% &	5444.827ms &	5445.809ms &	0.034ms	& 0.008ms &	944.464ms &	simple_go.js (line 835)\\
-\hline
-Add_Stone &	6010 &	1.1% &	2392.214ms	& 191461.602ms	& 31.857ms &	1.524ms &	1785.447ms &	simple_go.js (line 641)\\
-\hline
-\end{tabular}
+\begin{table}
+  \rotatebox{90}{
+    \begin{tabular}{ccccccccc} \toprule
+      Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	\midrule
+      (no name) &	404494 &	32.8 &	71255.203 &	147518.241 &	0.365 &	0.059 &	1650.149 &	simple\_go.js (line 407)\\
+      flood\_mark &	3838 &	15.94 &	34634.022 &	96713.52 &	25.199 &	0.206 &	1736.537 &	simple\_go.js (line 807)\\
+      iterate &	404495 &	8.75 &	19004.696 &	76254.33 &	0.189 &	0.011 &	1650.101 &	simple\_go.js (line 391)\\
+      (no name) &	735190 &	8.72 &	18941.197 &	18941.987 &	0.026 &	0.008 &	1572.648 &	simple\_go.js (line 819)\\
+      string\_as\_move &	242765 &	5.86 &	12724.373	& 12725.279 &	0.052 &	0.011 &	1629.186 &	simple\_go.js (line 192)\\
+      (no name) &	1041519 &	4.7 &	10209.513 &	57248.372 &	0.055 &	0.015 &	1649.869 &	simple\_go.js (line 416)\\
+      List\_Block\_Liberties &	7706 &	4.02 &	8727.331 &	21457.809 &	2.785 &	0.013 &	1630.607 &	simple\_go.js (line 863)\\
+      iterateOwnProperties &	14752 &	3.68 &	7985.019 &	9340.503 &	0.633 &	0.004 &	2.419 &	simple\_go.js (line 132)\\
+      update &	8742 &	3.18 &	6911.102 &	6911.93 &	0.791 &	0.004 &	1648.007 &	simple\_go.js (line 117)\\
+      Calculate\_Neighbour &	661 &	3.03 &	6593.032 &	60385.478 &	91.355 &	0.115 &	1679.244 &	simple\_go.js (line 829)\\
+      (no name) &	160414 &	2.51 &	5444.827 &	5445.809 &	0.034	& 0.008 &	944.464 &	simple\_go.js (line 835)\\
+      Add\_Stone &	6010 &	1.1 &	2392.214	& 191461.602	& 31.857 &	1.524 &	1785.447 &	simple\_go.js (line 641)\\ \bottomrule
+    \end{tabular}}
+  \caption{First 100 moves on a 19x19 board}
+  \label{table:first100}
+\end{table}
 
@@
 On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds.
-On an iPhone over AT&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
-
-Median move times in other go engines (gnugo especially)?
+On an iPhone over AT\&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
+%TODO: Median move times in other go engines (gnugo especially)?
 
 \chapter{Recommendations, especially for future work and unsolved issues}
-We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.
+We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo.
+Our anecdotal experience and those of others on the Web show that there is an order of magnitude difference between popular browsers for some JavaScript operations.
+However, we have not carefully studied this issue.%http://therealcrisp.xs4all.nl/blog/2006/12/09/string-performance-in-internet-explorer/
 

### 2007-12-19 paper/paper.tex
headings should be short

@@
 
-\chapter{Recommendations, especially for future work and unsolved issues}
+\chapter{Recommendations, future work, and unsolved issues}
 We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo.

### 2007-12-19 paper/paper.tex
label units

@@
     \begin{tabular}{ccccccccc} \toprule
-      Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	\midrule
+      Function & Calls & Percent & Own Time (ms) & Time (ms) & Avg (ms) & Min (ms) & Max (ms) & File\\	\midrule
       (no name) &	404494 &	32.8 &	71255.203 &	147518.241 &	0.365 &	0.059 &	1650.149 &	simple\_go.js (line 407)\\

### 2007-12-19 paper/paper.tex
align on decimal points in tables

@@
 \begin{table}
-  \rotatebox{90}{
-    \begin{tabular}{ccccccccc} \toprule
-      Function & Calls & Percent & Own Time (ms) & Time (ms) & Avg (ms) & Min (ms) & Max (ms) & File\\ \midrule
+  \rotatebox{90}{%  fc%         own       time      avg       min       max file
+    \begin{tabular}{lrD{.}{.}{2}D{.}{.}{3}D{.}{.}{3}D{.}{.}{3}D{.}{.}{3}D{.}{.}{3}l} \toprule
+      Function & Calls & \multicolumn{1}{c}{Percent} & \multicolumn{1}{c}{Own Time (ms)} & \multicolumn{1}{c}{Time (ms)} & \multicolumn{1}{c}{Avg (ms)} & \multicolumn{1}{c}{Min (ms)} & \multicolumn{1}{c}{Max (ms)} & File\\ \midrule
       (no name) & 109675 & 25.68 & 14310.594 &	36908.812 &	0.337 &	0.059	& 1593.542	& simple\_go.js (line 407)\\

### 2007-12-19 trunk/paper/paper.tex
refactor column alignment spec, and apply it to both tables

@@
 \tref{table:first50} shows the first 50 moves on a 19x19 board with human (black) versus crawler (white).
+\newcolumntype{M}{D{.}{.}{3}}
 \begin{table}
-  \rotatebox{90}{
-    \begin{tabular}{ccccccccc} \toprule
-      Function & Calls & Percent & Own Time (ms) & Time (ms) & Avg (ms) & Min (ms) & Max (ms) & File\\ \midrule
+  \rotatebox{90}{%  fc%         own       time      avg       min       max file
+    \begin{tabular}{lrD{.}{.}{2}MMMMMl} \toprule
+      Function & Calls & \multicolumn{1}{c}{Percent} & \multicolumn{1}{c}{Own Time (ms)} & \multicolumn{1}{c}{Time (ms)} & \multicolumn{1}{c}{Avg (ms)} & \multicolumn{1}{c}{Min (ms)} & \multicolumn{1}{c}{Max (ms)} & File\\ \midrule
       (no name) & 109675 & 25.68 & 14310.594 &	36908.812 &	0.337 &	0.059	& 1593.542	& simple\_go.js (line 407)\\
@@
   \rotatebox{90}{
-    \begin{tabular}{ccccccccc} \toprule
-      Function & Calls & Percent & Own Time & Time & Avg & Min & Max & File\\	\midrule
+    \begin{tabular}{lrD{.}{.}{2}MMMMMl} \toprule
+      Function & Calls & \multicolumn{1}{c}{Percent} & \multicolumn{1}{c}{Own Time (ms)} & \multicolumn{1}{c}{Time (ms)} & \multicolumn{1}{c}{Avg (ms)} & \multicolumn{1}{c}{Min (ms)} & \multicolumn{1}{c}{Max (ms)} & File\\ \midrule
       (no name) &	404494 &	32.8 &	71255.203 &	147518.241 &	0.365 &	0.059 &	1650.149 &	simple\_go.js (line 407)\\
@@
 On an iPhone over AT\&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
+%TODO: Median move times in other go engines (gnugo especially)?
 
-Median move times in other go engines (gnugo especially)?
-
-\chapter{Recommendations, especially for future work and unsolved issues}
+\chapter{Recommendations, future work, and unsolved issues}
 We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo.

### 2007-12-19 trunk/paper/paper.tex
case for JavaScript and Jaigo

@@
 
-Due to issues with the jaigo board representation, our measurements were limited to the first 100 moves.
+Due to issues with the Jaigo board representation, our measurements were limited to the first 100 moves.
 
-Regarding compactness, the total size of all jaigo files, including images, scripts, and html, was 102221 bytes for our initial release.
-The full implementation of the jaigo code is 66071 bytes \cite{SpeedReport}.
+Regarding compactness, the total size of all Jaigo files, including images, scripts, and html, was 102221 bytes for our initial release.
+The full implementation of the Jaigo code is 66071 bytes \cite{SpeedReport}.
 
-Regarding speed, we do not anticipate that the choice of processor and platform of any modern PC will have a significant change to the performance of jaigo.
+Regarding speed, we do not anticipate that the choice of processor and platform of any modern PC will have a significant change to the performance of Jaigo.
 
@@
 
-On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds.
-On an iPhone over AT\&T 2.5G EDGE with a full signal, the median jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
+On an iPhone over 54mbps 802.11g wifi with a full signal connected to a 3mbps cable modem, the median Jaigo load time was 5 seconds, with the initial load time being the longest measurement at 15 seconds.
+On an iPhone over AT\&T 2.5G EDGE with a full signal, the median Jaigo load time was 21 seconds, with a low of 17 seconds and a high of 30 seconds.
 %TODO: Median move times in other go engines (gnugo especially)?
@@
 \chapter{Recommendations, future work, and unsolved issues}
-We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo.
+We feel that the choice of browser is more important than the choice of computing platform for the performance of Jaigo.
 Our anecdotal experience and those of others on the Web show that there is an order of magnitude difference between popular browsers for some JavaScript operations.
@@
 
-The performance of $\alpha-\beta$ search pruning in javascript should be investigated. like SimpleGo v3, a 30kyu $\alpha$-$\beta$ pruned search move generator.
+The performance of $\alpha$-$\beta$ search pruning in JavaScript should be investigated. like SimpleGo v3, a 30kyu $\alpha$-$\beta$ pruned search move generator.
 
-There is much more optimization and correctness work to be done to the jaigo board and game representations.
-For example, the move speed of jaigo could be measured on many diverse computing platforms.
-Jaigo could be also used as a "widget", a popular way of packaging small javascript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.
+There is much more optimization and correctness work to be done to the Jaigo board and game representations.
+For example, the move speed of Jaigo could be measured on many diverse computing platforms.
+Jaigo could be also used as a ``widget'', a popular way of packaging small JavaScript and html applications into a format that is used by many different web sites and operating systems such as Windows Vista, Mac OS X, Facebook, Google, and others.
 
-There are many open issues being tracked by the jaigo project at this time. Addressing them may lead to better, more correct performance. We anticipate continuing work on this project.
+There are many open issues being tracked by the Jaigo project at this time. Addressing them may lead to better, more correct performance. We anticipate continuing work on this project.
 
